### 2012-10-15 notes.tex
Finished update equations in notes.

Isolating the terms of \eqref{eqn:composite} that involve $u_0$ gives
\begin{equation}
\mathcal{L}_{[u_0]} = K\sum_{j=1}^J\left(-\frac{1}{2\sigma_0^2}(\gamma_{1j}-u_0)^2\right)
\end{equation}

@@

{\bf Update Equation for $\sigma_0^2$ }

Isolating the terms of \eqref{eqn:composite} that involve $\sigma^2_0$ gives
\begin{equation}
\mathcal{L}_{[\sigma^2_0]} = -\frac{J}{2}\log(2\pi\sigma^2_0) -\frac{1}{2\sigma_0^2}\sum_{j=1}^J\left(\gamma_{2j} + (\gamma_{1j}-u_0)^2 \right)
\end{equation}

Taking the derivative with respect to $\sigma^2_0$ gives
\begin{equation}
\frac{\partial \mathcal{L}}{\partial \sigma^2_0} = -\frac{J}{2\sigma^2_0} +\frac{1}{\sigma^4_0}\sum_{j=1}^J\left(\gamma_{2j} + (\gamma_{1j}-u_0)^2\right)
\end{equation}

@@

{\bf Update Equation for $\alpha_{kj}$ and $\beta_{kj}$ }

Isolating terms that involve $\alpha_{kj}$ gives
\begin{multline}
\mathcal{L}_{[\alpha_{kj}]} = (u_0M_0 -1)(\psi(\alpha_{kj}) - \psi(\alpha_{kj}+\beta_{kj})) -((1-u_0)M_0-1)\psi(\alpha_{kj}+\beta_{kj}) \\
+ r_{kj}\psi(\alpha_{kj})-n_{kj}\psi(\alpha_{kj}+\beta_{kj}) \\
 - \log \Gamma (\alpha_{kj}+\beta_{kj}) + \log \Gamma(\alpha_{kj}) -(\alpha_{kj}-1)\psi(\alpha_{kj})+(\alpha_{kj}+\beta_{kj}-2)\psi(\alpha_{kj}+\beta_{kj})
\end{multline}

@@

A similar procedure holds for $\beta_{kj}$ and the resulting update equation bears symmetry to that for $\alpha_{kj}$
\begin{equation}
\beta_{kj} \leftarrow \psi_1(\beta_{kj})\left[(1-u_0)M_0-1+(n-r)-\beta_{kj}\right] - \psi_1(\alpha_{kj}+\beta_{kj})\left[M_0+n-(\alpha_{kj}+\beta_{kj})\right]
\end{equation}

{\bf Update Equation for $M_0$ }

Isolating terms that involve $M_0$ gives
\begin{multline}
\mathcal{L}_{[M_0]} = \log \Gamma(M_0) -\E_q\log \Gamma(\mu_jM_0) -\E_q\log \Gamma((1-\mu_j)M_0) +\\
+\gamma_{1j}M_0(\psi(\alpha_{kj})-\psi(\alpha_{kj}+\beta_{kj})) +(1-\gamma_{1j})M_0(\psi(\beta_{kj})-\psi(\alpha_{kj}+\beta_{kj}))
\end{multline}

{\bf Update Equation for $\gamma_{2j}$ }

Isolating terms that involve $\gamma_{2j}$ gives
\begin{equation}
\mathcal{L}_{[\gamma_{2j}]} = -\frac{K}{2\sigma_0^2}\gamma_{2j} -\E_q\log\Gamma(\mu_jM_0) - \E_q\log\Gamma((1-\mu_j)M_0)+\frac{1}{2}\log(2\pi e \gamma_{2j})
\end{equation}

>>>>

Isolating the terms of \eqref{eqn:composite} that involve $u_0$ gives
\begin{equation*}
\mathcal{L}_{[u_0]} = K\sum_{j=1}^J\left(-\frac{1}{2\sigma_0^2}(\gamma_{1j}-u_0)^2\right)
\end{equation*}

@@

%%%			sigma2 update			%%%
{\bf Update Equation for $\sigma_0^2$ }

Isolating the terms of \eqref{eqn:composite} that involve $\sigma^2_0$ gives
\begin{equation*}
\mathcal{L}_{[\sigma^2_0]} = -\frac{J}{2}\log(2\pi\sigma^2_0) -\frac{1}{2\sigma_0^2}\sum_{j=1}^J\left(\gamma_{2j} + (\gamma_{1j}-u_0)^2 \right)
\end{equation*}

Taking the derivative with respect to $\sigma^2_0$ gives
\begin{equation*}
\frac{\partial \mathcal{L}}{\partial \sigma^2_0} = -\frac{J}{2\sigma^2_0} +\frac{1}{\sigma^4_0}\sum_{j=1}^J\left(\gamma_{2j} + (\gamma_{1j}-u_0)^2\right)
\end{equation*}

@@

%%% 		M0 update 		%%%
{\bf Update Equation for $M_0$ }

Isolating terms that involve $M_0$ gives
\begin{multline*}
\mathcal{L}_{[M_0]} = \log \Gamma(M_0) -\E_q\log \Gamma(\mu_jM_0) -\E_q\log \Gamma((1-\mu_j)M_0) +\\
+\gamma_{1j}M_0(\psi(\alpha_{kj})-\psi(\alpha_{kj}+\beta_{kj})) +(1-\gamma_{1j})M_0(\psi(\beta_{kj})-\psi(\alpha_{kj}+\beta_{kj}))
\end{multline*}

Taking the derivative with respect to $M_0$ gives
\begin{multline*}
\frac{\partial \mathcal{L}}{\partial M_0} = \psi(M_0) -\frac{\partial}{\partial M_0}\E_q\log \Gamma(\mu_jM_0)
-\frac{\partial}{\partial M_0}\E_q\log \Gamma((1-\mu_j)M_0) \\
+\gamma_{1j}(\psi(\alpha_{kj})-\psi(\alpha_{kj}+\beta_{kj})) 
+(1-\gamma_{1j})(\psi(\beta_{kj})-\psi(\alpha_{kj}+\beta_{kj}))
\end{multline*}

The partial derivative can be taken inside of the integral which gives $\frac{\partial}{\partial M_0}\E_q\log \Gamma(\mu_jM_0) = \E_q \left[ \mu_j \psi(\mu_j M_0) \right]$ and a similar expression for the second term $\frac{\partial}{\partial M_0}\E_q\log \Gamma((1-\mu_j)M_0) = \E_q \left[(1- \mu_j) \psi((1-\mu_j) M_0) \right]$.

Plugging these simplifications into the partial with respect to $M_0$ gives
\begin{multline}
\frac{\partial \mathcal{L}}{\partial M_0} = \psi(M_0) - \E_q \left[ \mu_j \psi(\mu_j M_0) \right]
-\E_q \left[(1- \mu_j) \psi((1-\mu_j) M_0) \right] \\
+\gamma_{1j}(\psi(\alpha_{kj})-\psi(\alpha_{kj}+\beta_{kj})) 
+(1-\gamma_{1j})(\psi(\beta_{kj})-\psi(\alpha_{kj}+\beta_{kj}))
\end{multline}

%%%		alpha & beta update			%%%
{\bf Update Equation for $\alpha_{kj}$ and $\beta_{kj}$ }

Isolating terms that involve $\alpha_{kj}$ gives
\begin{multline*}
\mathcal{L}_{[\alpha_{kj}]} = (u_0M_0 -1)(\psi(\alpha_{kj}) - \psi(\alpha_{kj}+\beta_{kj})) -((1-u_0)M_0-1)\psi(\alpha_{kj}+\beta_{kj}) \\
+ r_{kj}\psi(\alpha_{kj})-n_{kj}\psi(\alpha_{kj}+\beta_{kj}) \\
 - \log \Gamma (\alpha_{kj}+\beta_{kj}) + \log \Gamma(\alpha_{kj}) -(\alpha_{kj}-1)\psi(\alpha_{kj})+(\alpha_{kj}+\beta_{kj}-2)\psi(\alpha_{kj}+\beta_{kj})
\end{multline*}

@@

A similar procedure holds for $\beta_{kj}$ and the resulting update equation bears symmetry to that for $\alpha_{kj}$,
\begin{equation}
 \frac{\partial \mathcal{L}}{\partial \beta_{kj}} = \psi_1(\beta_{kj}) \left[ (1-u_0)M_0-1+(n_{kj}-r_{kj})-\beta_{kj}\right] - \psi_1(\alpha_{kj}+\beta_{kj})(M_0+n-(\alpha_{kj}+\beta_{kj}))
\end{equation}

%%% 		gamma1 update 		%%%
{\bf Update equation for $\gamma_{1j}$}\label{sec:gamma1_update}

Isolating terms that involve $\gamma_{2j}$ gives
\begin{equation*}
\mathcal{L}_{[\gamma_{1j}]} = -\frac{K}{2\sigma_0^2}(\gamma_{1j}-u_0)^2 
+ \sum_{k=1}^K \gamma_{1j} M_0 ( \psi(\alpha_kj) - \psi(\alpha_{kj} + \beta_{kj}) )
- \sum_{k=1}^K \gamma_{1j} M_0 ( \psi(\beta_kj) - \psi(\alpha_{kj} + \beta_{kj}) )
\end{equation*}

Taking the derivative with respect to $\gamma_{1j}$, setting it equal to zero and solving for $\gamma_{1j}$ gives
\begin{equation}
\gamma_{1j} \leftarrow u_0 + M_0 \sigma_0^2 \frac{1}{K} \sum_{k=1}^K \psi(\alpha_{kj}) - \psi(\beta_{kj})
\end{equation}

%%% 		gamma2 update 		%%%
{\bf Update Equation for $\gamma_{2j}$ }\label{sec:gamma2_update}

Isolating terms that involve $\gamma_{2j}$ gives
\begin{equation*}
\mathcal{L}_{[\gamma_{2j}]} = -\frac{K}{2\sigma_0^2}\gamma_{2j} -\E_q\log\Gamma(\mu_jM_0) - \E_q\log\Gamma((1-\mu_j)M_0)+\frac{1}{2}\log(2\pi e \gamma_{2j})
\end{equation*}
